Beam-Helicity and Beam-Charge Asymmetries\\ Associated with Deeply Virtual Compton Scattering\\ on the Unpolarised Proton \underline{Paper Tag: PDVCS0607 (v3.0)} \underline{Drafting Committee: DC90 (\today)}

### 2012-03-08 paper.tex
Bibliography ordering was wrong...
Added gunars trivial changes. Also added one of the macros that have been left out of the repo...

\setvruler[15pt][1][1][3][1][30pt][30pt][-42pt]
\maketitle
\section{Introduction}
Generalised Parton Distributions (GPDs)
\cite{Mue94,Ji97,Rad97} encompass the familiar Parton
Distribution Functions (PDFs) and nucleon Form Factors (FFs) to provide a
comprehensive description of the structure of the nucleon. A thorough
description of the nucleon in terms of GPDs would allow the deduction
of the total angular momentum of partons in the nucleon, and the
construction of a longitudinal-momentum-dissected transverse spatial map of parton densities~\cite{Bur00}. The GPDs appear in experimental measurements in the form of complex-valued Compton Form Factors (CFFs), which are flavour-sums of convolutions of GPDs with hard scattering kernels. Constraints on these CFFs, and thus GPDs, can be obtained from measurements of exclusive leptoproduction processes. In particular, the exclusive leptoproduction of a single real photon from a nucleon or nucleus that remains intact ($e\,N\,\rightarrow\,e\,N\,\gamma$; see figure~\ref{spin}) is the simplest to describe and is the most widely-used reaction channel for such work (see \cite{Air01,Air06,Air08,Air09,Air10,Air10a,Air10b,Air11,Air11a,h101,h105,h107,h109,zeu03,zeu08,Maz07,Cam06,Ste01,Che06,Gir08,Gav09,Com10}).

@@

Events were selected if having exactly one photon and one lepton
track detected within the acceptance of the spectrometer.
The event selection is subject to the kinematic constraints 1\,GeV$^{2}$ $<$
Q$^{2}$ $<$ 10\,GeV$^{2}$, 0.03 $<$ $x_{\textrm{B}}$ $<$ 0.35,
$-t < 0.7$\,GeV$^2$, $W^{2}$ $>$
9\,GeV$^{2}$ and $\nu$ $<$ 22\,GeV, where $W$ is the invariant mass of the
$\gamma^{*}p$ system and $\nu$ is the energy of the virtual photon in the target
rest frame. The polar angle between the directions of the virtual and real photons was required to
be within the limits 5~mrad $<$
$\theta_{\gamma^{*}\gamma}$ $<$ 45~mrad.

@@

\section{Background Corrections and Systematic Uncertainties for the
  2006-2007 Data}
The extracted asymmetry amplitudes are subject to systematic uncertainties that result from a combination of background processes, shifts in the missing-mass distributions, and various detector and binning effects determined in the same manner as used in refs.~\cite{Air08,Air09} in order to maintain consistency with the results published in ref.~\cite{Air09} and therefore facilitate the combination of the two data sets.

The contribution to the uncertainties on the amplitude measurements
arising from background in the data from neutral meson
production is predominantly due to the failure to identify one of
  the two photons from the decay of these neutral mesons. The procedure for estimating the corresponding
uncertainty on the measured amplitude values is described in detail in
refs.~\cite{Air08,Air09}. Each measurement in the $-t$, $x_{\textrm{B}}$ and
  $Q^{2}$ projections has this uncertainty estimated at the
  centre of the relevant kinematic bin and included as part of the
  total systematic uncertainty.

@@

The first and second harmonics of $\mathcal{A}_{\textrm{LU,I}}$, which are sensitive to the interference term in the scattering amplitude, are shown in the first and third rows of figure~\ref{bsa_xbjrange}. The leading-twist amplitude $A_{\textrm{LU,I}}^{\sin\phi}$ has the largest magnitude of any of the amplitudes when extracted in a single bin from the entire data set. This amplitude shows no strong dependence on $-t$, $x_{\textrm{B}}$ or $Q^{2}$, implying a strong dependence at smaller values of $-t$ as the amplitude must approach zero as $-t$ approaches its minimum value because of the dependence of the amplitude on the factor $K$ defined in eq.~\ref{eq:K}. The $A_{\textrm{LU,I}}^{\sin\phi}$ amplitude is sensitive to the imaginary part of the CFF $\mathcal{H}$ and thereby can constrain GPD $\textit{H}$. The $A_{\textrm{LU,DVCS}}^{\sin\phi}$ asymmetry is shown in the second row of figure~\ref{bsa_xbjrange}. Both the $A_{\textrm{LU,DVCS}}^{\sin\phi}$ asymmetry amplitude and the $A_{\textrm{LU,I}}^{\sin(2\phi)}$ asymmetry amplitude are compatible with zero, and neither asymmetry amplitude shows any dependence on $-t$, $x_{\textrm{B}}$ or $Q^{2}$.

@@

In order to provide more detailed information that can be used in future fits, in particular for the determination of the entanglement of the skewness and $-t$ dependences of GPDs, the amplitudes already presented in figures~\ref{bsa_xbjrange} and~\ref{bca_xbjrange} are shown as a function of $-t$ for three different ranges of $x_{\textrm{B}}$ in figures~\ref{bsa_xbjrange2} and~\ref{bca_xbjrange2}. These figures represent the kinematic dependences of the amplitudes in a less-correlated manner than the one-dimensional projections: within experimental uncertainty, there is no evidence of a correlation between the $-t$ and $x_{\textrm{B}}$ dependences for any of the amplitudes. The results from this paper will be made available in the Durham Database. The results will also be made available in the same 4-bin format as used in previous analyses at H{\sc ermes}~\cite{Air08,Air09,Air10b}.

\section{Summary}
Beam-helicity and beam-charge asymmetries in the azimuthal
distribution of real photons from leptoproduction on an unpolarised
hydrogen target have been presented. These asymmetries were extracted
from an unpolarised hydrogen data set taken during the 2006 and 2007
operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were
extracted previously from hydrogen data obtained during the 1996-2005
experimental period as described in ref.~\cite{Air09}. A comparison of
the amplitudes extracted from these independent data sets has shown
that they are compatible and the asymmetry amplitudes can therefore be
extracted from the complete 1996-2007 data set. The asymmetry
amplitudes extracted from the complete data set are the most
statistically precise DVCS measurements presented by H{\sc
  ermes}. There is a strong signal in the first harmonic of the
interference contribution to the beam-helicity asymmetry. There are
non-zero amplitudes in the zeroth and first harmonics of the
beam-charge asymmetry. All asymmetry amplitudes related to higher
harmonics are consistent with zero. The results from the complete data
set are compared to calculations from ongoing work to fit GPD models to experimental data. All asymmetry amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.

@@

Figure~\ref{pic:covmat} shows the covariance matrix for the 13 parameter fit for all amplitudes extracted in a single bin across the entire kinematic range. The covariance matrix for the fit in each of the kinematic bins presented in figures~\ref{bsa_xbjrange2} and~\ref{bca_xbjrange2} will be made available in the Durham database.

@@

\caption{The covariance matrix results for the asymmetries extracted in a single bin across the whole kinematic range.}

@@

X.~Ji, \emph{Deeply virtual Compton scattering}
\\Phys. Rev. D \textbf{55} (1997) 7114 [arxiv:hep-ph/9609381].

@@

@@

\bibitem{Maz07}
JLab Hall A Collaboration, M. Mazouz et al., \emph{Deeply virtual Compton scattering off the neutron}
\\Phys. Rev. Lett. {\bf 99} (2007) 242501 [arxiv:0709.0450].

\bibitem{Cam06}
JLab Hall A Collaboration,  C. Camacho et al., \emph{Scaling Tests of the Cross Section for Deeply Virtual Compton Scattering}
\\Phys. Rev. Lett. \textbf{97} (2006) 262002 [arxiv:nucl-ex/0607029].

\bibitem{Ste01}
JLab C{\sc las} Collaboration, S. Stepanyan et al., \emph{Observation of exclusive DVCS in polarized electron beam asymmetry measurements}
\\Phys. Rev. Lett. {\bf 87} (2001) 182002 [arxiv:hep-ex/0107043].

\bibitem{Che06}
JLab C{\sc las} Collaboration, S. Chen et al., \emph{Measurement of Deeply Virtual Compton Scattering with a Polarized Proton Target}
\\Phys. Rev. Lett. {\bf 97} (2006) 072002 [arxiv:hep-ex/0605012].

\bibitem{Gir08}
JLab C{\sc las} Collaboration, F. X. Girod et al., \emph{Measurement of Deeply Virtual Compton Scattering Beam-Spin Asymmetries}
\\Phys. Rev. Lett. \textbf{100} (2008) 162002 [arxiv:0711.4805].

\bibitem{Gav09}
JLab C{\sc las} Collaboration, G. Gavalian et al., \emph{Beam Spin Asymmetries in DVCS with CLAS with 4.8GeV}
\\Phys. Rev. C {\bf 80} (2009) 035206 [arxiv:0812.2950].

@@

\bibitem{Kum09}
K. Kumericki and D. M\"uller, \emph{Deeply virtual Compton scattering at small $x_B$ and the access to the GPD H}
\\Nucl. Phys. B {\bf 841} (2010) 1 [arxiv:0904.0458].

@@

\bibitem{Kum07}
K. Kumericki, D. M\"uller and K. Passek-Kumericki, \emph{Towards a fitting procedure for deeply virtual Compton scattering at next-to-leading order and beyond}
\\Nucl. Phys. B \textbf{794} (2008) 244 [arxiv:hep-ph/0703179].

\bibitem{Kum08}
K. Kumericki, D. M\"uller and K. Passek-Kumericki, \emph{Sum rules and dualities for generalized parton distributions: is there a holographic principle?}
\\Eur. Phys. J. C \textbf{58} (2008) 193 [arxiv:0805.0152].

>>>>

\maketitle
\section{Introduction}
Generalised Parton Distributions (GPDs)
\cite{Mue94,Ji97,Rad97} encompass the familiar Parton
Distribution Functions (PDFs) and nucleon Form Factors (FFs) to provide a
comprehensive description of the structure of the nucleon. A thorough
description of the nucleon in terms of GPDs would allow the deduction
of the total angular momentum of partons in the nucleon, and the
construction of a longitudinal-momentum-dissected transverse spatial map of parton densities~\cite{Bur00}. The GPDs appear in experimental measurements in the form of complex-valued Compton Form Factors (CFFs), which are flavour-sums of convolutions of GPDs with hard scattering kernels. Constraints on these CFFs, and thus GPDs, can be obtained from measurements of exclusive leptoproduction processes. In particular, the exclusive leptoproduction of a single real photon from a nucleon or nucleus that remains intact ($e\,N\,\rightarrow\,e\,N\,\gamma$; see figure~\ref{spin}) is the simplest to describe and is the most widely-used reaction channel for such work (see \cite{Ji97b,Air01,Air06,Air08,Air09,Air10,Air10a,Air10b,Air11,Air11a,h101,h105,h107,h109,zeu03,zeu08,Maz07,Cam06,Ste01,Che06,Gir08,Gav09,Com10}).

@@

Events in the 2006-2007 data set were selected if having exactly one
\red{lepton} track detected within the acceptance of the spectrometer and \red{at
  least one photon depositing $>5$\,GeV in the electromagnetic
  calorimeter. The latter selection criterion differs from the photon
  selection criterion used for the 1996-2005 data set as an intermittent
  hardware fault in 2006-2007 can cause spurious noise signals in the
  calorimeter that are misinterpreted as very low energy photons.}
The event selection is subject to the kinematic constraints 1\,GeV$^{2}$ $<$
Q$^{2}$ $<$ 10\,GeV$^{2}$, 0.03 $<$ $x_{\textrm{B}}$ $<$ 0.35,
$-t < 0.7$\,GeV$^2$, $W^{2}$ $>$
9\,GeV$^{2}$ and $\nu$ $<$ 22\,GeV, where $W$ is the invariant mass of the
$\gamma^{*}p$ system and $\nu$ is the energy of the virtual photon in the target
rest frame. The polar angle between the directions of the virtual and real photons was required to
be within the limits 5~mrad $<$
$\theta_{\gamma^{*}\gamma}$ $<$ 45~mrad.

@@

\section{Background Corrections and Systematic Uncertainties for the
  2006-2007 Data}
The extracted asymmetry amplitudes are subject to systematic uncertainties
that result from a combination of background processes, shifts in the
missing-mass distributions, and various detector and binning effects
determined in the same manner as used in refs.~\cite{Air08,Air09} in order to
maintain consistency with the results published in ref.~\cite{Air09} and
therefore facilitate the combination of the two data sets. \red{No systematic
  uncertainty is assigned from the intermittent fault in the calorimeter
  mentioned in section 2; the percentage of events affected by the issue is
  small for the 2006-2007 data sample and completely negligible in the context
of the combined data sets.}

The contribution to the uncertainties on the amplitude measurements
arising from background in the data from neutral meson
production is predominantly due to the failure to identify one of
  the two photons from the decay of these neutral mesons. \red{The
    $\mathcal{A}^{\textrm{DVCS}}_{\textrm{LU}}$ amplitudes are corrected for
  inclfuences from semi-inclusive pion production and an uncertainty is
  assigned to the affected amplitudes. A further uncertainty is assigned to
  these amplitudes for the influence of photons produced in decays of
  exclusively produced pions.} The procedure for estimating the corresponding
uncertainty on the measured amplitude values is described in detail in
refs.~\cite{Air08,Air09}. Each measurement in the $-t$, $x_{\textrm{B}}$ and
  $Q^{2}$ projections has this uncertainty estimated at the
  centre of the relevant kinematic bin and included as part of the
  total systematic uncertainty. \red{No correction is applied or uncertainty
    assigned to the amplitudes for the
    $\mathcal{A}_{\textrm{LU}}^{\textrm{I}}$ or $\mathcal{A}_{\textrm{C}}$
    asymmetries because influences from meson production are expected to
    vanish when considering the difference between beam charges.}

@@

The first and second harmonics of $\mathcal{A}^{\textrm{I}}_{\textrm{LU}}$, which are sensitive to the interference term in the scattering amplitude, are shown in the first and third rows of figure~\ref{bsa_xbjrange}. The leading-twist amplitude $A_{\textrm{LU,I}}^{\sin\phi}$ has the largest magnitude of any of the amplitudes when extracted in a single bin from the entire data set. This amplitude shows no strong dependence on $-t$, $x_{\textrm{B}}$ or $Q^{2}$, implying a strong dependence at smaller values of $-t$ as the amplitude must approach zero as $-t$ approaches its minimum value because of the dependence of the amplitude on the factor $K$ defined in eq.~\ref{eq:K}. The $A_{\textrm{LU,I}}^{\sin\phi}$ amplitude is sensitive to the imaginary part of the CFF $\mathcal{H}$ and thereby can constrain GPD $\textit{H}$. The $A_{\textrm{LU,DVCS}}^{\sin\phi}$ asymmetry is shown in the second row of figure~\ref{bsa_xbjrange}. Both the $A_{\textrm{LU,DVCS}}^{\sin\phi}$ asymmetry amplitude and the $A_{\textrm{LU,I}}^{\sin(2\phi)}$ asymmetry amplitude are compatible with zero, and neither asymmetry amplitude shows any dependence on $-t$, $x_{\textrm{B}}$ or $Q^{2}$.

@@

In order to provide more detailed information that can be used in future fits,
in particular for the determination of the entanglement of the skewness and
$-t$ dependences of GPDs, the amplitudes already presented in
figures~\ref{bsa_xbjrange} and~\ref{bca_xbjrange} are shown as a function of
$-t$ for three different ranges of $x_{\textrm{B}}$ in
figures~\ref{bsa_xbjrange2} and~\ref{bca_xbjrange2}. These figures represent
the kinematic dependences of the amplitudes in a less-correlated manner than
the one-dimensional projections: within experimental uncertainty, there is no
evidence of a correlation between the $-t$ and $x_{\textrm{B}}$ dependences
for any of the amplitudes.

The results from this paper will be made available in the Durham Database. The results will also be made available in the same 4-bin format as used in previous analyses at H{\sc ermes}~\cite{Air08,Air09,Air10b}.

\section{Summary}
Beam-helicity and beam-charge asymmetries in the azimuthal
distribution of real photons from \red{hard exclusive} leptoproduction on an unpolarised
hydrogen target have been presented. These asymmetries were extracted
from an unpolarised hydrogen data set taken during the 2006 and 2007
operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were
extracted previously from hydrogen data obtained during the 1996-2005
experimental period as described in ref.~\cite{Air09}. A comparison of
the amplitudes extracted from these independent data sets has shown
that they are compatible and the asymmetry amplitudes can therefore be
extracted from the complete 1996-2007 data set. The asymmetry
amplitudes extracted from the complete data set are the most
statistically precise DVCS measurements presented by H{\sc
  ermes}. There is a strong signal in the first harmonic of the
interference contribution to the beam-helicity asymmetry. There are
non-zero amplitudes in the zeroth and first harmonics of the
beam-charge asymmetry. All asymmetry amplitudes related to higher
harmonics are consistent with zero. The results from the complete data
set are compared to calculations from ongoing work to fit GPD models to experimental data. All asymmetry amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.

@@

Figure~\ref{pic:covmat} shows the covariance matrix for the 13 parameter fit for all amplitudes extracted in a single bin across the entire kinematic range. The covariance matrix for the fit in each of the kinematic bins presented in figures~\ref{bsa_xbjrange} to~\ref{bca_xbjrange2} will be made available in the Durham database.

@@

\caption{The covariance matrix results for the asymmetries extracted in a
  single bin across the whole kinematic range. \red{The size of the symbols in
the chart reflect the magnitude of the corresponding correlation. Closed
(open) symbols represent positive (negative) correlations.}}

@@

@@

\bibitem{Ji97b}
X.~Ji, \emph{Deeply virtual Compton scattering}
\\Phys. Rev. D \textbf{55} (1997) 7114 [arxiv:hep-ph/9609381].

@@

\bibitem{Maz07}
Jefferson Lab Hall A Collaboration, M. Mazouz et al., \emph{Deeply virtual Compton scattering off the neutron}
\\Phys. Rev. Lett. {\bf 99} (2007) 242501 [arxiv:0709.0450].

\bibitem{Cam06}
Jefferson Lab Hall A Collaboration,  C. Camacho et al., \emph{Scaling Tests of the Cross Section for Deeply Virtual Compton Scattering}
\\Phys. Rev. Lett. \textbf{97} (2006) 262002 [arxiv:nucl-ex/0607029].

\bibitem{Ste01}
C{\sc las} Collaboration, S. Stepanyan et al., \emph{Observation of exclusive DVCS in polarized electron beam asymmetry measurements}
\\Phys. Rev. Lett. {\bf 87} (2001) 182002 [arxiv:hep-ex/0107043].

\bibitem{Che06}
C{\sc las} Collaboration, S. Chen et al., \emph{Measurement of Deeply Virtual Compton Scattering with a Polarized Proton Target}
\\Phys. Rev. Lett. {\bf 97} (2006) 072002 [arxiv:hep-ex/0605012].

\bibitem{Gir08}
C{\sc las} Collaboration, F. X. Girod et al., \emph{Measurement of Deeply Virtual Compton Scattering Beam-Spin Asymmetries}
\\Phys. Rev. Lett. \textbf{100} (2008) 162002 [arxiv:0711.4805].

\bibitem{Gav09}
C{\sc las} Collaboration, G. Gavalian et al., \emph{Beam Spin Asymmetries in DVCS with CLAS with 4.8GeV}
\\Phys. Rev. C {\bf 80} (2009) 035206 [arxiv:0812.2950].

@@

\bibitem{Kum09}
K. Kumeri\v{c}ki and D. M\"uller, \emph{Deeply virtual Compton scattering at small $x_B$ and the access to the GPD H}
\\Nucl. Phys. B {\bf 841} (2010) 1 [arxiv:0904.0458].

@@

\bibitem{Kum07}
K. Kumeri\v{c}ki, D. M\"uller and K. Passek-Kumeri\v{c}ki, \emph{Towards a fitting procedure for deeply virtual Compton scattering at next-to-leading order and beyond}
\\Nucl. Phys. B \textbf{794} (2008) 244 [arxiv:hep-ph/0703179].

\bibitem{Kum08}
K. Kumeri\v{c}ki, D. M\"uller and K. Passek-Kumeri\v{c}ki, \emph{Sum rules and dualities for generalized parton distributions: is there a holographic principle?}
\\Eur. Phys. J. C \textbf{58} (2008) 193 [arxiv:0805.0152].
